Variational Approximations for Transfer in Reinforcement Learning

### 2018-05-15 NIPS 2018/main.tex
app c

\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\arginf}{\operatornamewithlimits{arginf}}
\newcommand{\argsup}{\operatornamewithlimits{argsup}}
\newcommand{\mm}{\operatornamewithlimits{mm}}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}
\title{Variational Approximations for Transfer in Reinforcement Learning}

@@

Recent advancements have allowed reinforcement learning (RL)~\cite{sutton1998reinforcement} to achieve impressive results in a wide variety of complex tasks, ranging from Atari~\cite{mnih2015human} through the game of Go~\cite{silver2016mastering} to the control of sophisticated robotics systems~\cite{kober2009policy,lillicrap2015continuous,levine2016end}. The main limitation is that these RL algorithms still require an enormous amount of experience samples before successfully learning such complicated tasks. One of the most promising solutions to reduce the need of samples is transfer learning, which focuses on reusing past knowledge available to the agent in order to reduce the sample-complexity for learning new tasks. In the typical settings of transfer in RL~\cite{taylor2009transfer}, the agent is assumed to have already solved a set of \textit{source tasks} generated from some unknown distribution. Then, given a \textit{target task} (which is drawn from the same distribution, or a slightly different one), the agent can rely on knowledge from the source tasks to speed up the learning process. This reuse of knowledge constitutes a significant advantage over plain RL, where the agent learns each new task from scratch independently of any previous learning experience. Several algorithms have been proposed in the literature to transfer different elements involved in the learning process: experience samples~\cite{lazaric2008transfer,taylor2008transferring}, policies/options~\cite{fernandez2006probabilistic,konidaris2007building}, rewards~\cite{konidaris2006autonomous}, features~\cite{barreto2017successor}, parameters~\cite{doshi2016hidden,killian2017robust}, and so on. We refer the reader to~\cite{taylor2009transfer,lazaric2012transfer} for a thorough survey on transfer in RL.

%\paragraph*{Exploration-based introduction}
%One of the most relevant problems in this context is how to efficiently explore the target task based on knowledge from the source tasks. Intuitively, assuming the tasks under consideration share some similarities due to the common distribution, much better exploration strategies than uninformed ones (e.g., $\epsilon$-greedy) can be adopted for quickly learning the target task. Among the appealing approaches for this problem we find Bayesian methods (e.g.,~\cite{}), which are able to model the uncertainty over the current task based on previous knowledge and drive exploration so that this uncertainty is reduced as quickly as possible. Similarly, model-based algorithms (e.g.,~\cite{}) typically transfer samples to improve their estimates of the task model and adopt classic count-based exploration to drive the agent towards regions where such estimates are more uncertain. However, all these approaches either require strong assumptions (for example, on the distribution involved in Bayesian methods) or do not scale well to large problems. This greatly limits their practical applicability.

%In this work, we tackle such limitations by considering a more general approach. Similarly to~\cite{}, we assume tasks to share similarities in their value functions and use the given source tasks to learn the distribution over such functions. Then, we use this distribution as a prior for learning the target task and we propose and efficient variational approximation of the corresponding posterior. Leveraging on recent ideas from randomized value functions (\cite{}), we design a Thompson sampling-based algorithm which efficiently explores the target task by repeatedly sampling from the posterior and acting greedily w.r.t. (with respect to) the sampled value function. We show that our approach is very general, in the sense that it does not require any specific choice of function approximator or prior/posterior distribution models. \todo[inline]{Maybe mention something about theory/experiments here}

Under the assumption that tasks follow a specific distribution, an intuitive choice for designing a transfer algorithm is to attempt at characterizing the uncertainty over the target task. Then, an ideal algorithm would leverage prior knowledge from the source tasks to interact with the target task to reduce the uncertainty as quickly as possible. This simple intuition makes Bayesian \todo{probabilistic? distributional?} methods appealing approaches for transfer in RL, and many previous works have been proposed in this direction. In~\cite{wilson2007multi}, the authors assume tasks share similarities in their dynamics and rewards and propose a hierarchical Bayesian model for the distribution of these two elements. Similarly, in~\cite{lazaric2010bayesian}, the authors assume that tasks are similar in their value functions and design a different hierarchical Bayesian model for transferring such information. More recently,~\cite{doshi2016hidden}, and its extension~\cite{killian2017robust}, consider tasks whose dynamics are governed by some hidden parameters, and propose efficient Bayesian models for quickly learning such parameters in new tasks. However, most of these algorithms require specific, and sometimes restrictive, assumptions (e.g., on the distributions involved or the function approximators adopted), which might limit their practical applicability. The importance of having transfer algorithms that alleviate the need for strong assumptions and that easily adapt to different contexts motivates us to take a more general approach.

Similarly to~\cite{lazaric2010bayesian}, we assume tasks to share similarities in their value functions and use the given source tasks to learn a distribution over such functions. Then, we use this distribution as a prior for learning the target task and we propose a variational approximation of the corresponding posterior that is computationally efficient. Leveraging on recent ideas from randomized value functions~\cite{osband2014generalization}, we design a Thompson Sampling-based algorithm which efficiently explores the target task by repeatedly sampling from the posterior and acting greedily w.r.t. (with respect to) the sampled value function. We show that our approach is very general, in the sense that it can work with any parametric function approximator and with any prior/posterior distribution models (in this paper we focus on the Gaussian and Gaussian mixture models). In addition to the algorithmic contribution, we give also a theoretical contribution by providing a finite-sample analysis of our approach and an experimental contribution showing its empirical performance on four domains with increasing level of difficulty.

@@

We consider a distribution $\mathcal{D}$ over tasks, where each task $\mathcal{M}_\tau$ is modeled as a discounted Markov Decision Process (MDP). We define an MDP as a tuple $\mathcal{M}_\tau = \langle \mathcal{S},\mathcal{A},\mathcal{P}_\tau,\mathcal{R}_\tau,p_0,\gamma\rangle$, where $\mathcal{S}$ is the state-space, $\mathcal{A}$ is a finite set of actions, $\mathcal{P}_\tau(\cdot | s,a)$ is the distribution of the next state $s'$ given that action $a$ is taken in state $s$, $\mathcal{R}_\tau: \mathcal{S}\times\mathcal{A}\rightarrow\mathbb{R}$ is the reward function, $p_0$ is the initial-state distribution, and $\gamma\in [0,1)$ is the discount factor. We assume the reward function to be uniformly bounded by a constant $R_{max}>0$. A deterministic policy $\pi : \mathcal{S} \rightarrow \mathcal{A}$ is a mapping from states to actions. At the beginning of each episode of interaction, the initial state $s_0$ is drawn from $p_0$. Then, the agent takes the action $a_0 = \pi(s_0)$, receives a reward $\mathcal{R}_\tau(s_0,a_0)$, transitions to the next state $s_1 \sim \mathcal{P}(\cdot | s_0,a_0)$, and the process is repeated. The goal is to find the policy maximizing the long-term return over a possibly infinite horizon: $\max_{\pi}J(\pi)\triangleq\mathbb{E}_{\mathcal{M}_\tau,\pi}[\sum_{t=0}^{\infty} \gamma^t \mathcal{R}_\tau(s_t,a_t)]$. To this end, we define the optimal value function of task $\mathcal{M}_\tau$, $Q_\tau^*(s,a)$, as the expected return obtained by taking action $a$ in state $s$ and following an optimal policy thereafter. Then, an optimal policy $\pi_\tau^*$ is a policy that is greedy with respect to the optimal value function, i.e., $\pi_\tau^*(s) = \argmax_a Q_\tau^*(s,a)$ for all states $s$. It can be shown (e.g.,~\cite{puterman1994markov}) that $Q_\tau^*$ is the unique fixed-point of the optimal Bellman operator $T_\tau$ defined by $T_\tau Q(s,a) = \mathcal{R}_\tau(s,a) + \gamma\mathbb{E}_{s' \sim\mathcal{P}_\tau}[\max_{a'}Q(s',a')]$ for any value function $Q$. From now on, we adopt the term $Q$-function to denote any plausible value function, i.e., any function $Q : \mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ uniformly bounded by $\frac{R_{max}}{1-\gamma}$. In the following, to avoid cluttering the notation, we will drop the subscript $\tau$ when there is no ambiguity.

We consider a parametric family of $Q$-functions, $\mathcal{Q} = \left\{ Q_{\bm{w}} : \mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R} \mid \bm{w}\in\mathbb{R}^d\right\}$, and we assume each function in $\mathcal{Q}$ to be uniformly bounded by $\frac{R_{max}}{1-\gamma}$. When learning the optimal value function, a quantity of interest is how close a given function $Q_{\bm{w}}$ is to the fixed-point of the Bellman operator. A possible measure is its Bellman error (or Bellman residual), defined by $B_{\bm{w}} \triangleq TQ_{\bm{w}} - Q_{\bm{w}}$. Notice that $Q_{\bm{w}}$ is optimal if and only if $B_{\bm{w}}(s,a) = 0$ for all $s,a$. If we assume the existence of a distribution $\nu$ over $\mathcal{S}\times\mathcal{A}$, a sound objective is to directly minimize the squared Bellman error of $Q_{\bm{w}}$ under $\nu$, denoted by $\norm{B_{\bm{w}}}_{\nu}^2$. Unfortunately, it is well-known that an unbiased estimator of this quantity requires two independent samples of the next state $s'$ for each $s,a$ (e.g.,~\cite{maillard2010finite}). In practice, the Bellman error is typically replaced by the TD error $b(\bm{w})$, which approximates the former using a single transition sample $\langle s,a,s',r \rangle$, $b(\bm{w}) = r + \gamma\max_{a'}Q_{\bm{w}}(s',a') - Q_{\bm{w}}(s,a)$. Finally, given a dataset $D = \langle s_i,a_i,r_i,s_i' \rangle_{i=1}^{N}$ of $N$ samples, the squared TD error is computed as $\norm{B_{\bm{w}}}_{D}^2 = \frac{1}{N}\sum_{i=1}^N (r_i + \gamma \max_{a'} Q_{\bm{w}}(s_i',a') - Q_{\bm{w}}(s_i,a_i))^2 = \frac{1}{N}\sum_{i=1}^N b_i(\bm{w})^2$. Whenever the distinction is clear from the context, with slight abuse of terminology, we refer to the squared Bellman error and squared TD error as Bellman error and TD error, respectively.

%\subsection{Variational Inference}
%
%When working with Bayesian approaches, the posterior distribution of hidden variables $\bm{w} \in \mathbb{R}^K$ given data $D$,
%\begin{equation}\label{eq:bayes}
%p(\bm{w} | D) = \frac{p(D | \bm{w})p(\bm{w})}{p(D)} = \frac{p(D | \bm{w})p(\bm{w})}{\int_{\bm{w}} p(D | \bm{w})p(\bm{w})},
%\end{equation}
%is typically intractable for many models of interest (e.g., when working with deep neural networks) due to difficulties in computing the integral of Eq. \eqref{eq:bayes}. The main intuition behind variational inference~\cite{} \todo{CITE} is to approximate the intractable posterior $p(\bm{w} | D)$ with a simpler distribution $q_{\bm{\xi}}(\bm{w})$. The latter is chosen in a parametric family, with variational parameters $\bm{\xi}$, as the minimizer of the Kullback-Leibler (KL) divergence w.r.t. $p$:
%\begin{equation}
%\min_{\bm{\xi}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w} \mid D)\right)
%\end{equation}
%It is well-known that minimizing the KL divergence is equivalent to maximizing the so-called \textit{evidence lower bound} (ELBO), which is defined as:
%\begin{equation}
%\text{ELBO}(\bm{\xi}) = \mathbb{E}_{\bm{w} \sim q_{\bm{\xi}}}\left[ \log p(D | \bm{w}) \right] - KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right)
%\end{equation}
%Intuitively, the best approximation is the one that maximizes the expected log-likelihood of the data, while minimizing the KL divergenge w.r.t. the prior $p(\bm{w})$.

@@

A first important question that we need to answer is whether replacing max with mellow-max in the Bellman operator constitutes a strong approximation or not. It has been proved \cite{asadi2017alternative} that the mellow Bellman operator is a non-expansion under the $L_{\infty}$-norm and, thus, has a unique fixed-point.  However, how such fixed-point differs from the one of the optimal Bellman operator remains an open question. Since mellow-max monotonically converges to max as $\kappa \rightarrow \infty$, it would be desirable if the corresponding operator also monotonically converged to the optimal one. We confirm that this property actually holds in the following theorem.

@@

The proof is provided in App. \ref{}. As expected, $\wt{Q}$ converges to $Q^*$ exponentially fast when either $\kappa$ (equivalently, $\beta_{\kappa}$) increases, or the action gaps are enlarged. Notice that this result is of interest even outside our specific settings.

The second question that we need to answer is whether we can provide any guarantee on our algorithm's performance when given limited data. To address this point, we restrict to the case of linear approximators with Gaussian prior and posterior. We assume Alg. \ref{alg:transfer} is used with a finite dataset and provide a finite-sample analysis bounding the distance between the fixed-point of the mellow Bellman operator and $Q$-functions sampled from the variational distribution minimizing the objective \eqref{eq:elbo}. Our main results is given in the following theorem.

\begin{restatable}{theorem}{thg} \label{th:main-gaussian}
Fix a target task $\tau$ a let $\wt{Q}$ be the fixed-point of the corresponding mellow Bellman operator. Assume linearly parameterized value functions $Q_{\bm{w}}(s,a)=\bm{w}^T\bm{\phi}(s,a)$ with bounded weights $\bm{w}\leq w_{max}$ and uniformly bounded features $\bm{\phi}(s,a)\leq \phi_{\max}$. Consider the Gaussian version of Alg. \ref{alg:transfer} with prior $p(\bm{w}) =\mathcal{N}(\bm{\mu}_p,\bm{\Sigma}_p)$ and denote by $(\wh{\bm{\mu}},\wh{\bm{\Sigma}})$ the variational parameter minimizing the objective of Eq. \ref{eq:elbo} on a dataset $D$ of $N$ samples. Let $\nu$ be a probability measure over $\mathcal{S}\times\mathcal{A}$ and $\bm{w}^* = \arginf_{\bm{w}}\norm{\wt{B}(\bm{w})}_{\nu}^2$. Define $\upsilon(\bm{w}^*) \triangleq \mathbb{E}_{\mathcal{N}(\bm{w}^*,\bm{I})}\left[ v(\bm{w}) \right]$, with $v(\bm{w}) \triangleq \mathbb{E}_{\nu}\left[Var_{\mathcal{P}}\left[b(\bm{w})\right]\right]$. Then, there exist constants $c_1,c_2,c_3$ such that, with probability at least $1-2\delta$ over the choice of weights $\bm{w} \sim \mathcal{N}(\wh{\bm{\mu}},\wh{\bm{\Sigma}})$ and dataset $D$:
\begin{equation}
\resizebox{0.93\textwidth}{!}{$
\norm{Q_{\bm{w}} - \wt{Q}}_{\nu}^2 \leq \frac{1}{(1-\gamma)\delta}\left(2\norm{\wt{B}(\bm{w}^{*})}_{\nu}^2 + \frac{\upsilon(\bm{w}^*) + c_3\sqrt{\log\frac{2}{\delta}}}{\sqrt{N}} + \frac{c_2 + \lambda\norm{\bm{w}^* - \bm{\mu}_p}_{\bm{\Sigma}_p^{-1}}}{N} + \frac{c_1}{N^2}\right)$}
\end{equation}
\end{restatable}
We refer the reader to App. \ref{} for the proof and a specific definition of the three constants. Four main terms constitute our bound: the approximation error due to the limited hypothesis space (first term), the variance (second term), the distance to the prior (third term), and a constant term decaying very fast (as $\mathcal{O}(\frac{1}{N^2})$). Intuitively, our bound demonstrates that, if one starts very close the prior, the only error incurred, besides the irreducible approximation error, is due to the variance of the estimates, and, thus, the algorithm is expected to achieve good performance rather quickly. Furthermore, as $N\rightarrow\infty$ the only error term remaining is the irreducible approximation error. Notice that the variance term $\upsilon(\bm{w}^*)$ is due to the fact that we minimize a biased estimator of the Bellman error. As mentioned in the preliminaries, if double sampling of the next state is possible (e.g., in simulation), such term would disappear. Finally, we want to point out that the proof of this theorem relies on a very general result (Lemma \ref{lemma:general-bound} in App. \ref{}) that does not require any specific choice of distribution of approximator. Leveraging on such results, it is straightforward to provide a finite-sample analysis of the mixture version of our algorithm. For the sake of completeness, we report the derivation in App. \ref{}.

@@

In this section, we provide an experimental evaluation of our approach in four different domains with increasing level of difficulty. In all experiments, we compare our Gaussian variational transfer algorithm (GVT) and the version using a $c$-component mixture of Gaussians ($c$-MGVT) to plain no-transfer RL (NT) with $\epsilon$-greedy exploration. To the best of our knowledge, no existing transfer algorithm is directly comparable to our approach from an experimental perspective. Thus, we provide a discussion of related works in the next section. \todo{Any better motivation?}

@@

We consider an agent navigating in the rooms environment depicted in Fig. \ref{}. The agent starts in the bottom-left corner and must move from one room to another to reach the goal position in the top-right corner. The rooms are connected by small doors whose positions are unknown to the agent. The state-space is modeled as a $10 \times 10$ continuous grid, while the action-space is the set of $4$ movement directions (up, right, down, left). After each action, the agent moves by $1$ in the chosen direction and the final position is corrupted by Gaussian noise with $0.2$ standard deviation. In case the agent hits a wall, its position remains unchanged. The reward is $1$ when reaching the goal (after which the process terminates) and $0$ otherwise, while the discount factor is $\gamma = 0.99$. In this experiment, we consider linearly parameterized $Q$-functions with $121$ equally-spaced radial basis features.

Since the agent does not know where the doors are located in advance and receives only very sparse feedback, it must efficiently explore the environment to figure out (i) their positions, and (ii) how to reach the goal. While this might be a complicated problem for plain RL, our transfer algorithm should be able to quickly figure out the door positions. In fact, notice that, although different, the optimal $Q$-functions for all tasks share some similarities. For example, once the agent has passed the last door before the goal, the $Q$-values are exactly the same in all tasks. This does not hold for positions nearby the start state. However, it is clear that there should be a preference over actions up and right, rather than down and left (which are worse in all tasks).\todo{Good candidate to be removed}

@@

In order to prove that our guesses are correct, we generate a set of $50$ source tasks for the three-room environment of Fig. \ref{} by sampling both door positions uniformly in the allowed space, and solve all of them by directly minimizing the TD error as presented in Sec. \ref{sec:td}. Then, we use our algorithms to transfer from $10$ source tasks sampled from the previously generated set. The average return over the last $50$ learning episodes as a function of the number of iterations is shown in Fig. \ref{}. Each curve is the result of $20$ independent runs, each resampling the target and source tasks, with $95\%$ confidence intervals. Further details on the parameters adopted in this experiment are given in App. \ref{}. As expected, the no-transfer (NT) algorithm fails at learning the task in so few iterations due to the limited exploration provided by an $\epsilon$-greedy policy. On the other hand, all our algorithms achieve a significant speed-up and are able to converge to the optimal performance in few iterations, with GVT being slightly slower. Interestingly, we notice that there is no advantage in adopting more than $1$ component for the posterior in MGVT. This is intuitive since, when the algorithm quickly figures out which task is being solved, it will move all components in the same direction.

To better understand the differences between GVT and MGVT, we now consider transferring from a slightly different distribution than the one from which target tasks are drawn. We generate again $50$ source tasks but this time with the bottom door fixed at the center and the other one moving. Then, we repeat the previous experiment, allowing both doors to move when sampling target tasks. The results are shown in Fig. \ref{}. Interestingly, MGVT seems almost unaffected by this change, proving that is has sufficient representation power to generalize to slightly different task distributions. The same does not hold for GVT, which now is not able to solve many of the sampled target tasks, as can be noticed from the higher variance. This proves again that assuming Gaussian distributions can pose severe limitations in our transfer settings.

>>>>

\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\mm}{\operatornamewithlimits{mm}}
\newcommand{\wt}[1]{\widetilde{#1}}
\newcommand{\wh}[1]{\widehat{#1}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}
\title{Variational Approximations for Transfer in Reinforcement Learning}

@@

Recent advancements have allowed reinforcement learning (RL) \cite{sutton1998reinforcement} to achieve impressive results in a wide variety of complex tasks, ranging from Atari \cite{mnih2015human}, to the game of Go \cite{silver2016mastering}, to the control of sophisticated robotics systems \cite{kober2009policy,lillicrap2015continuous,levine2016end}. The main limitation is that RL algorithms still require an enormous amount of experience samples before successfully learning such complicated tasks. One of the most promising solutions is transfer learning, which focuses on reusing past knowledge available to the agent in order to reduce the sample-complexity for learning new tasks. In the typical settings of transfer in RL \cite{taylor2009transfer}, the agent is assumed to have already solved a set of \textit{source tasks} generated from some unknown distribution. Then, given a \textit{target task} drawn from the same distribution, or a slightly different one, the agent can rely on knowledge from the source tasks to speed up the learning process. This constitutes a significant advantage over plain RL, where the agent learns each new task from scratch independently of previous learning experience. Several algorithms have been proposed in the literature to transfer experience samples \cite{lazaric2008transfer,taylor2008transferring}, policies/options \cite{fernandez2006probabilistic,konidaris2007building}, rewards \cite{konidaris2006autonomous}, features \cite{barreto2017successor}, parameters \cite{doshi2016hidden,killian2017robust}, and so on. We refer the reader to \cite{taylor2009transfer} for a thorough survey on transfer in RL.

%\paragraph*{Exploration-based introduction}
%One of the most relevant problems in this context is how to efficiently explore the target task based on knowledge from the source tasks. Intuitively, assuming the tasks under consideration share some similarities due to the common distribution, much better exploration strategies than uninformed ones (e.g., $\epsilon$-greedy) can be adopted for quickly learning the target task. Among the appealing approaches for this problem we find Bayesian methods (e.g., \cite{}), which are able to model the uncertainty over the current task based on previous knowledge and drive exploration so that this uncertainty is reduced as quickly as possible. Similarly, model-based algorithms (e.g., \cite{}) typically transfer samples to improve their estimates of the task model and adopt classic count-based exploration to drive the agent towards regions where such estimates are more uncertain. However, all these approaches either require strong assumptions (for example, on the distribution involved in Bayesian methods) or do not scale well to large problems. This greatly limits their practical applicability.

%In this work, we tackle such limitations by considering a more general approach. Similarly to \cite{}, we assume tasks to share similarities in their value functions and use the given source tasks to learn the distribution over such functions. Then, we use this distribution as a prior for learning the target task and we propose and efficient variational approximation of the corresponding posterior. Leveraging on recent ideas from randomized value functions (\cite{}), we design a Thompson sampling-based algorithm which efficiently explores the target task by repeatedly sampling from the posterior and acting greedily w.r.t. (with respect to) the sampled value function. We show that our approach is very general, in the sense that it does not require any specific choice of function approximator or prior/posterior distribution models. \todo[inline]{Maybe mention something about theory/experiments here}

Under the assumption that tasks follow a certain distribution, an intuitive choice for designing a transfer algorithm is to attempt at characterizing the uncertainty over the target task. Then, an ideal algorithm would leverage prior knowledge from the source tasks to interact with the target task in such a way that this uncertainty is reduced as quickly as possible. This simple intuition makes Bayesian \todo{probabilistic? distributional?} methods appealing approaches for transfer in RL, and many previous works have been proposed in this direction. \cite{wilson2007multi} assume tasks share similarities in their dynamics and rewards and propose a hierarchical Bayesian model for the distribution of these two elements. Similarly, \cite{lazaric2010bayesian} assume tasks are similar in their value functions and design a different hierarchical Bayesian model for transferring such information. More recently, \cite{doshi2016hidden}, and its extension \cite{killian2017robust}, consider tasks whose dynamics are governed by some hidden parameters, and propose efficient Bayesian models for quickly learning such parameters in new tasks. However, all these algorithms require specific, and sometimes restrictive, assumptions (e.g., on the distributions involved or the function approximators adopted), which might limit their practical applicability. The importance of having transfer algorithms that alleviate the need of strong assumptions and that can be easily adapted to different contexts motivates us to take a more general approach.

Similarly to \cite{lazaric2010bayesian}, we assume tasks to share similarities in their value functions and use the given source tasks to learn a distribution over such functions. Then, we use this distribution as a prior for learning the target task and we propose and efficient variational approximation of the corresponding posterior. Leveraging on recent ideas from randomized value functions \cite{osband2014generalization}, we design a Thompson Sampling-based algorithm which efficiently explores the target task by repeatedly sampling from the posterior and acting greedily w.r.t. (with respect to) the sampled value function. We show that our approach is very general, in the sense that it does not require any specific choice of function approximator or prior/posterior distribution models. We propose a finite-sample analysis that theoretically validates our approach, while empirically evaluating its performance on three domains with increasing level of difficulty.

@@

We consider a distribution $\mathcal{D}$ over tasks, where each task is modeled as a discounted Markov Decision Process (MDP). We define an MDP as a tuple $\mathcal{M} = \langle \mathcal{S},\mathcal{A},\mathcal{P},\mathcal{R},p_0,\gamma\rangle$\todo{Should we provide a more transfer-oriented definition?}, where $\mathcal{S}$ is the state-space, $\mathcal{A}$ is a finite set of actions, $\mathcal{P}(\cdot | s,a)$ is the distribution of the next state $s'$ given that action $a$ is taken in state $s$, $\mathcal{R}: \mathcal{S}\times\mathcal{A}\rightarrow\mathbb{R}$ is the reward function, $p_0$ is the initial-state distribution, and $\gamma\in [0,1)$ is the discount factor. We assume the reward function to be uniformly bounded by a constant $R_{max}>0$. A deterministic policy $\pi : \mathcal{S} \rightarrow \mathcal{A}$ is a mapping from states to actions. At the beginning of each episode of interaction, the initial state $s_0$ is drawn from $p_0$. Then, the agent takes the action $a_0 = \pi(s_0)$, receives a reward $\mathcal{R}(s_0,a_0)$, transitions to the next state $s_1 \sim \mathcal{P}(\cdot | s_0,a_0)$, and the process is repeated. The goal is to find the policy maximizing the long-term return over a possibly infinite horizon: $\max_{\pi}J(\pi)\triangleq\mathbb{E}[\sum_{t=0}^{\infty} \gamma^t r_t \mid \mathcal{M},\pi]$. To this end, we define the optimal value function $Q^*(s,a)$ as the expected return obtained by taking action $a$ in state $s$ and following an optimal policy thereafter. Then, an optimal policy $\pi^*$ is a policy that is greedy with respect to the optimal value function, i.e., $\pi^*(s) = \argmax_a Q^*(s,a)$ for all states $s$. It can be shown (e.g., \cite{puterman1994markov}) that $Q^*$ is the unique fixed-point of the optimal Bellman operator $T$ defined by $TQ(s,a) = \mathcal{R}(s,a) + \gamma\mathbb{E}_{\mathcal{P}}[\max_{a'}Q(s',a')]$ for any value function $Q$. From now on, we adopt the term $Q$-function to denote any plausible value function, i.e., any function $Q : \mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ uniformly bounded by $\frac{R_{max}}{1-\gamma}$.

We consider a parametric family of $Q$-functions, $\mathcal{Q} = \left\{ Q_{\bm{w}} : \mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R} \mid \bm{w}\in\mathbb{R}^K\right\}$, and we assume each function in $\mathcal{Q}$ to be uniformly bounded by $\frac{R_{max}}{1-\gamma}$. When learning the optimal value function, a quantity of interest is how close a given function $Q_{\bm{w}}$ is to the fixed-point of the Bellman operator. A possible measure is its Bellman error (or Bellman residual), defined by $B(\bm{w}) \triangleq TQ_{\bm{w}} - Q_{\bm{w}}$. Notice that $Q_{\bm{w}}$ is optimal if, and only if, $B(\bm{w})(s,a) = 0$ for all $s,a$. If we assume the existence of a distribution $\nu$ over $\mathcal{S}\times\mathcal{A}$, an appealing objective is to directly minimize the squared Bellman error of $Q_{\bm{w}}$ under $\nu$, denoted by $\norm{B(\bm{w})}_{\nu}^2$\todo{Should we define norms before?}. Unfortunately, it is well-known that an unbiased estimator of this quanitity requires two independent samples of the next state $s'$ for each $s,a$ (e.g., \cite{maillard2010finite}). In practice, the Bellman error is typically replaced by the TD error $b(\bm{w})$, which approximates the former using a single transition sample, $b(\bm{w}) = r + \gamma\max_{a'}Q_{\bm{w}}(s',a') - Q_{\bm{w}}(s,a)$. Finally, given a dataset $D = \langle s_i,a_i,r_i,s_i' \rangle_{i=1}^{N}$ of $N$ samples, the squared TD error is computed as $\norm{B(\bm{w})}_{D}^2 = \frac{1}{N}\sum_{i=1}^N (r_i + \gamma \max_{a'} Q_{\bm{w}}(s_i',a') - Q_{\bm{w}}(s_i,a_i))^2 = \frac{1}{N}\sum_{i=1}^N b_i(\bm{w})^2$. Whenever the distinction is clear from the context, with slight abuse of terminology, we refer to the squared Bellman error and squared TD error as Bellman error and TD error, respectively.

%\subsection{Variational Inference}
%
%When working with Bayesian approaches, the posterior distribution of hidden variables $\bm{w} \in \mathbb{R}^K$ given data $D$,
%\begin{equation}\label{eq:bayes}
%p(\bm{w} | D) = \frac{p(D | \bm{w})p(\bm{w})}{p(D)} = \frac{p(D | \bm{w})p(\bm{w})}{\int_{\bm{w}} p(D | \bm{w})p(\bm{w})},
%\end{equation}
%is typically intractable for many models of interest (e.g., when working with deep neural networks) due to difficulties in computing the integral of Eq. \eqref{eq:bayes}. The main intuition behind variational inference \cite{} \todo{CITE} is to approximate the intractable posterior $p(\bm{w} | D)$ with a simpler distribution $q_{\bm{\xi}}(\bm{w})$. The latter is chosen in a parametric family, with variational parameters $\bm{\xi}$, as the minimizer of the Kullback-Leibler (KL) divergence w.r.t. $p$:
%\begin{equation}
%\min_{\bm{\xi}} KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w} \mid D)\right)
%\end{equation}
%It is well-known that minimizing the KL divergence is equivalent to maximizing the so-called \textit{evidence lower bound} (ELBO), which is defined as:
%\begin{equation}
%\text{ELBO}(\bm{\xi}) = \mathbb{E}_{\bm{w} \sim q_{\bm{\xi}}}\left[ \log p(D | \bm{w}) \right] - KL\left(q_{\bm{\xi}}(\bm{w})\ ||\ p(\bm{w})\right)
%\end{equation}
%Intuitively, the best approximation is the one that maximizes the expected log-likelihood of the data, while minimizing the KL divergenge w.r.t. the prior $p(\bm{w})$.

@@

In this section, we theoretically analyze our variational transfer algorithm...

A first important question that we need to answer is whether replacing max with mellow-max in the Bellman operator constitutes a strong approximation or not. It has been proved \cite{} \todo{Cite MM} that the mellow Bellman operator is a contraction under the $L_{\infty}$-norm and, thus, has a unique fixed-point.  However, how such fixed-point differs from the one of the optimal Bellman operator remains an open question. Since mellow-max monotonically converges to max as $\kappa \rightarrow \infty$, it would be desirable if the corresponding operator also monotonically converged to the optimal one. We confirm that this property actually holds in the following theorem.

@@

@@

In this section, we provide an experimental evaluation of our approach in three different domains with increasing level of difficulty. In all experiments, we compare our Gaussian variational transfer algorithm (GVT) and the version using a $c$-component mixture of Gaussians ($c$-MGVT) to plain no-transfer RL (NT) with $\epsilon$-greedy exploration. To the best of our knowledge, no existing transfer algorithm is directly comparable to our approach from an experimental perspective. \todo{Any better motivation?}

@@

We consider an agent navigating in the rooms environment depicted in Fig. \ref{}. The agent starts in the bottom-left corner and must move from one room to another to reach the goal position in the top-right corner. The rooms are connected by small doors whose positions are unknown to the agent. The state-space is modeled as a $10 \times 10$ continuous grid, while the action-space is the set of $4$ movement directions (up, right, down, left). After each action, the agent moves by $1$ in the chosen direction and the final position is corrupted by Gaussian noise with $0.2$ standard deviation. In case the agent hits a wall, its position remains unchanged. The reward is $1$ when reaching the goal (after which the process terminates) and $0$ otherwise, while the discount factor is $\gamma = 0.99$. We consider a distribution over tasks in which doors have a fixed width of $1$, while their positions are sampled uniformly in $[0.5,9.5]$. Since the agent does not know where the doors are located in advance and receives only very sparse feedback, it must efficiently explore the environment to figure out (i) their positions, and (ii) how to reach the goal. While this might be a complicated problem for plain RL, our transfer algorithm should be able to quickly figure out the door positions. In fact, notice that, although different, the optimal $Q$-functions for all tasks share some similarities. For example, once the agent has passed the last door before the goal, the $Q$-values are exactly the same in all tasks. This does not hold for positions nearby the start state. However, it is clear that there should be a preference over actions up and right, rather than down and left (which are worse in all tasks). Thus, we expect our algorithm to efficiently explore any target task.

@@

In order to prove that our guesses are correct, we generate a set of $50$ source tasks for the three-room environment of Fig. \ref{} by sampling both door positions uniformly, and solve all of them by directly minimizing the TD error as presented in Sec. \ref{sec:td}. In order to make sure that their solutions are accurate enough, we allow sampling the initial state uniformly in the environment and run until convergence. Then, we use our algorithms to transfer from $10$ source tasks sampled from the previously generated set. The average return over the last $50$ learning episodes as a function of the number of iterations is shown in Fig. \ref{}. Each curve is the result of $20$ independent runs, each resampling the target and source tasks. $95\%$ confidence intervals are shown. Further details on the parameters adopted in this experiment are given in App. \ref{}. As expected, the no-transfer (NT) algorithm fails at learning the task in so few iterations due to the limited exploration provided by an $\epsilon$-greedy policy. On the other hand, all our algorithms achieve a significant speed-up and are able to converge to the optimal performance in few iterations, with GVT being slightly slower. Interestingly, we notice that there is no advantage in adopting more than $1$ component for the posterior in MGVT. This is intuitive since, when the algorithm quickly figures out which task is being solved, it will move all components in the same direction.

To better understand the differences between GVT and MGVT, we now consider transferring from a slightly different distribution than the one from which target tasks are drawn. We generate again $50$ source tasks but this time with the bottom door fixed in the center and the other one moving. Then, we repeat the previous experiment, allowing both doors to move when sampling target tasks. The results are shown in Fig. \ref{}. Interestingly, MGVT seems almost unaffected by this change, proving that is has sufficient representation power to generalize to slightly different task distributions. The same does not hold for GVT, which now is not able to solve many of the sampled target tasks,thus the very high variance. This proves again that assuming Gaussian distributions can pose severe limitations in our transfer settings.
